### 2025-02-21 Priors table.tex
Updates from Overleaf

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{soul}
\bibliographystyle{plain}
\usepackage{amsmath}

\begin{document}

\begin{table}[h]
    \centering
    \begin{tabular}{lll}
         & \textbf{Description} & \textbf{Prior} \\
&Data & \\
\hline
$K$ & known DNA concentration from qPCR standards in copies/$\mu$L & - \\
$Z$ & qPCR amplification (yes=1; no=0) & - \\
$Y$ & qPCR cycle threshold (Ct) & - \\
$R$ & metabarcoding number of reads & - \\
$T$ & total number of reads for all S species & - \\
$N_{PCR}$ & number of metabarcoding PCR cycles & - \\
$L$ & known DNA concentration from mock commuinties in copies/$\mu$L & - \\
$V$ & volume filtered and aliquote dilution offset & -\\
$I$ & aliquote dilution factor & -\\

&&\\
&State processes&\\
\hline
$E_{s\neq H}$ & unknown DNA concentration in copies/$\mu$L & $\mathcal{N}(0,5)$\\
$D$ & unknown DNA concentration in copies/L & - \\
$F$ & smoothed DNA concentration in copies/L & - \\
&&\\

&Parameters&\\
\hline
$\psi$& overall mean hake eDNA concentration (copies/$\mu$L) & $\mathcal{N}(2,2)$\\
$\epsilon_{(x,y,d)}$& spatial and depth specific random effect & $\mathcal{N}(0,\sigma_D)$\\
$\sigma_D= (0, +\infty)$& standard deviation of spatial and depth specific random effect & $\mathcal{N}(0,3)$\\

$\delta$& biological replicate random effect & $\mathcal{N}(0,\tau)$\\
$\tau = (0, +\infty)$& standard deviation of biological replicate random effect & $\mathcal{N}(0,0.2)$\\

$\phi$& intercept of the qPCR probability of detection relationship & $\mathcal{N}(3,1)$\\
$\beta0$& intercept of the linear relation between the mean Ct values&$\mathcal{N}(39,3)$\\
&and eDNA concentration (K and E) & \\
$\beta1$& slope of the linear relation between the mean Ct values and &$\mathcal{N}(-1.4,0.05)$\\
&eDNA concentration (K and E) & \\
$\gamma0$& intercept of the linear relation between the standard deviation&$\mathcal{N}(0,1)$\\
&of Ct values and eDNA concentration (K and E) & \\
$\gamma1$& intercept of the linear relation between the standard deviation &$\mathcal{N}(0,1)$\\
&of Ct values and eDNA concentration (K and E) & \\
$\alpha$& species amplification efficiency & fixed\\
$\zeta$& ethanol wash effect & fixed\\
&&\\
&Subscripts&\\
\hline
$s$& species ($H$ = Hake) & -\\
$x$& northing & -\\
$y$& easting & -\\
$d$& depth & -\\
$i$& biological replicate & -\\
$r$& technical replicate & -\\
$p$& qPCR plate & -\\
$m$& mock community sample & -\\
$j$& qPCR standard aliquote sample & -\\

    \end{tabular}
    \caption{Data, state processes, parameters, and subscripts employed in the joint Bayesian model and their prior distributions}
    \label{tab:filter_error}
\end{table}

\end{document}